Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-18 doc/GMP2015/s01_introduction.tex
White space fixes.

% \newcommand{\meeraheader}[2]{\medskip\noindent\underbar{{\sl (#1) #2}}}

@@

@@

    \item Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
    materials that behave like assemblies of 2D particles under
    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}
\end{enumerate}
%
In order to study structural and mechanical properties
of a material layer, it is natural to model a material layer as a solution or
realization of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D
\dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined
formally in Section \ref{sec:prelim}) can be
used to understand or design material layers (their solutions) with
desired properties.

\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
qusecs,  many of which are \dfn{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}),
\dfn{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

@@

\medskip\noindent
\header{(v) Configuration Spaces of Under-constrained Systems.}
For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed \uncited\uncited to complete them into isostatic systems
\uncited\uncited and to find paths within the connected components
\uncited\uncited of standard Cartesian configuration spaces. Most of the
algorithms with formal guarantees leverage Cayley configuration space
theory \uncited\uncited to characterize subclasses of graphs and
additional constraints that control combinatorial explosion, and
provide faithful bijective representation of connected components and
paths. These algorithms have decreasing efficiency as the subclass of
systems gets bigger, with highest efficiency for underlying partial
2-tree graphs (alternately, tree-width 2, series- parallel, $K_4$
minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
graphs with low Cayley complexity (which include common linkages such
as the Strandbeest, Limacon and Cardioid), and decreased efficiency
for general 1-dof tree-decomposable graphs. While software suites
exist \uncited\uncited, no such formal algorithms and guarantees are known
for non-tree-decomposable systems.

\subsection{Contributions and Organization}
\label{sec:cont}

>>>>

@@

\noindent
\textbf{Note on Scope:} In this manuscript we only study finite 2D structures.
\dfn{Self-similarity} refers to the result of finitely many levels of hierarchy
or subdivision in an iterated scheme to generate self-similar structures.

@@

    \item Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and
    materials that behave like assemblies of 2D particles under
    non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}
\end{enumerate}
%
In order to study structural and mechanical properties
of a material layer, it is natural to model a material layer as a solution or
realization of a geometric constraint system of appropriate types of geometric
primitives, under metric or algebraic constraints. Such 2D
\dfn{qusecs} (quasi-uniform or self-similar constraint systems, defined
formally in Section \ref{sec:prelim}) can be
used to understand or design material layers (their solutions) with
desired properties.
%
\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
We now briefly survey existing techniques for studying 2D
qusecs,  many of which are \dfn{bar-joint} systems
(Examples 1, 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, \ref{sec:recomb}),
\dfn{body-hyperpin} systems (Example 4,5, see Section
\ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example
3, see Section \ref{sec:pinnedline}). The limitations of these
techniques directly motivate the contributions of this paper.

@@

\medskip\noindent
\header{(v) Configuration Spaces of Under-constrained Systems.}
For under-constrained 2D bar-joint and body-hyperpin qusecs obtained
from various subclasses of tree-decomposable systems, algorithms have
been developed \uncited\uncited to complete them into isostatic systems
\uncited\uncited and to find paths within the connected components
\uncited\uncited of standard Cartesian configuration spaces. Most of the
algorithms with formal guarantees leverage Cayley configuration space
theory \uncited\uncited to characterize subclasses of graphs and
additional constraints that control combinatorial explosion, and
provide faithful bijective representation of connected components and
paths. These algorithms have decreasing efficiency as the subclass of
systems gets bigger, with highest efficiency for underlying partial
2-tree graphs (alternately called, tree-width 2, series-parallel, $K_4$
minor avoiding), moderate efficiency for 1 degree- of-freedom (dof)
graphs with low Cayley complexity (which include common linkages such
as the Strandbeest, Limacon and Cardioid), and decreased efficiency
for general 1-dof tree-decomposable graphs. While software suites
exist \uncited\uncited, no such formal algorithms and guarantees are known
for non-tree-decomposable systems.
%
\subsection{Contributions and Organization}
\label{sec:cont}

### 2014-12-18 doc/GMP2015/s02_preliminaries.tex
White space fixes.

\section{Preliminaries}
\label{sec:prelim}
%
\subsection{Basics}
%
In this paper, a \dfn{geometric constraint system} is a multivariate
polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$,
representing constraints with parameters $\delta$ between geometric
primitives  in $\mathbb{R}^2$ represented collectively as $x\in
\mathbb{R}^n$.
%
When the type of constraint (system) is fixed, the system is simply
represented as $(G,\delta)$, where $G$ is the underlying constraint
(hyper)graph $G = (V,E)$ with the vertices $V$ representing the
geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$
representing the constraints, each with an associated parameter
$\delta$.
%
For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a
graph $G=(V,E)$ with fixed length bars as edges, i.e. $\delta: E
\rightarrow \mathbb{R}$; this represents the distance constraint
system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where
$x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.
%
\noindent
\note Geometric constraint systems can also have
inequalities in addition to equations, where the parameters in $\delta$ are small intervals of
values rather than exact values.

@@

In this paper, a \dfn{qusec} is any \dfn{independent} geometric
constraint system of one of the 3 types mentioned above.
\medskip\noindent
\note In the remainder of this section and the next three
sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint
qusecs and graphs. Relevant formal analogies for the other 2 types of
qusecs and (hyper)graphs are given in the subsequent 2 sections on
Applications.

@@

%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{over-constrained} if it is not independent.
%
%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
% \begin{definition}
%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}

% \begin{definition}
%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}

% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
%
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \uncited {\sl DRP}$(G)$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex
        set of $C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
%
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.

    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan

%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
%an isostatic vertex-maximal proper subgraph.
    %Note that nodes will be referred to interchangeably as ``the node that
    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}
%
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
    For a given graph, there could be exponentially many  DR-plans and
    even optimal DR-plans in the size of the graph . A complete DR-
    plan is unique but may not be (is usually not) optimal. DR-plans
    of self-similar graphs are self-similar. More than one node (leaf)
    in a DR-plan forest may represent the same subgraph (vertex) of
    $G$.
\end{remark}

>>>>

\section{Preliminaries}
\label{sec:prelim}
%
\subsection{Basics}
%
In this paper, a \dfn{geometric constraint system} is a multivariate
polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$,
representing constraints with parameters $\delta$ between geometric
primitives  in $\mathbb{R}^2$ represented collectively as $x\in
\mathbb{R}^n$.
%
When the type of constraint (system) is fixed, the system is simply
represented as $(G,\delta)$, where $G$ is the underlying constraint
(hyper)graph $G = (V,E)$ with the vertices $V$ representing the
geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$
representing the constraints, each with an associated parameter
$\delta$.
%
For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a
graph $G=(V,E)$ with fixed length bars as edges, i.e. $\delta: E
\rightarrow \mathbb{R}$; this represents the distance constraint
system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where
$x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.

\noindent
\note Geometric constraint systems can also have
inequalities in addition to equations, where the parameters in $\delta$ are small intervals of
values rather than exact values.

@@

In this paper, a \dfn{qusec} is any \dfn{independent} geometric
constraint system of one of the 3 types mentioned above.

\medskip\noindent
\note In the remainder of this section and the next three
sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint
qusecs and graphs. Relevant formal analogies for the other 2 types of
qusecs and (hyper)graphs are given in the subsequent 2 sections on
Applications.

@@

%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%    \item \textbf{over-constrained} if it is not independent.
%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
% \begin{definition}
%     A graph is \textbf{over-constrained} if, given constant $k$, there exists some
%     % induced
%     subgraph $S\subseteq G$ such that $d(S) > k$.
% \end{definition}
% \begin{definition}
%     A graph is \textbf{well-constrained} if, given constant $k$, $d(G) = k$ and for all
%     % induced
%     subgraphs $S\subseteq G$, (1) $d(S)\leq k$, or (2) given a set of trivial graphs $T$, $S$ is isomorphic to one of the graphs in $T$ (i.e.\ $S$ is trivial).
% \end{definition}
% \begin{definition}
%     A \textbf{trivial} graph is ill-defined in the general case. The only strict requirements are: (1) it must be an over-constrained graph and (2) all of its subgraphs are also trivial.
%     \todo{Maybe leave the following out until it's needed later?}
%     In the familiar geometric cases of $d$-dimensional space, all vertex weights are $d$, all edge weights are $1$, and constant $k= -{{d+1}\choose{2}}$. Trivial graphs for 2D would be a single vertex. Trivial graphs in 3D would be a vertex and an edge (2 vertices with an edge between). Etc. These trivial graphs capture the notion of the rotational symmetry that exists in geometric spaces.
% \end{definition}
\begin{definition}\label{def:drp}
    The \dfn{decomposition-recombination \mbox{(DR-)} plan} of graph
    $G$, \uncited {\sl DRP}$(G)$ (see Figure \ref{fig:c2c3ofk33s}, is defined as a forest that has the
    following properties:
    \begin{enumerate}
        \item Each node represents/contains/is a rigid vertex-induced
        subgraph of $G$.
        % \item For a node, $C$, that is  a non-trivial graph, its $N$
        % children, $C_1, \ldots, C_N$, are rigid vertex-maximal proper
        % subgraphs of $C$.
        \item The vertex set of $\bigcup_{i=1}^N{C_i}$ is the vertex
        set of $C$.
        \item A leaf node is a single vertex.
A \dfn{trivial} graph is empty or a single vertex.
        \item A root node is a maximal rigid subgraph of $G$.
    \end{enumerate}
    % It can also be described recursively as: the root is $G$, its children are the trivial subgraphs and the DR-plans of its well-constrained vertex-maximal proper subgraphs whose union is $G$ itself.
    A DR-plan is \dfn{complete} if it satisfies an additional rule:
        \item For a node $C$ that is  a non-trivial graph, its $N$
        children, $C_1, \ldots, C_N$, are all of the rigid vertex-
        maximal proper subgraphs of $C$. This makes Rule 2 implicit,
        and in fact, $C$ contains all the edges in the union of its
        children as well.
    A DR-plan is \dfn{optimal} if it minimizes the maximum fan-in over all nodes in the tree.
    The maximum  fan-in is called the \dfn{size} of the DR-plan.
%In general, $C$ is the graph in an arbitrary node in $CompleteDRP(G)$.
%$C_i$ is the $i^{\text{th}}$ child of $C$. It is implied that $C_i$ is
%an isostatic vertex-maximal proper subgraph.
    %Note that nodes will be referred to interchangeably as ``the node that
    %represents or contains the (sub)graph $C$'' and as simply ``$C$''.
\end{definition}
% \todo{This was labeled as both begin(remark) and end(observation), what was your intent?}
\begin{remark}
    For a given graph, there could be exponentially many  DR-plans and
    even optimal DR-plans in the size of the graph . A complete DR-
    plan is unique but may not be (is usually not) optimal. DR-plans
    of self-similar graphs are self-similar. More than one node (leaf)
    in a DR-plan forest may represent the same subgraph (vertex) of
    $G$.
\end{remark}
